---
title: "Passing Control Args"
author: "Steven M. Mortimer"
date: "2019-06-08"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 4
    keep_md: true
vignette: >
  %\VignetteIndexEntry{Passing Control Args}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo = FALSE}
NOT_CRAN <- identical(tolower(Sys.getenv("NOT_CRAN")), "true")
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  purl = NOT_CRAN,
  eval = NOT_CRAN
)
```

If you're inserting records from R you may want to turn off the assignment rules 
or even bypass duplicate rules and alerts to save records. Beginning in Version 0.1.3 of 
the **salesforcer** package many functions have a `control` argument that will allow 
you to fine tune the behavior of calls to the Salesforce APIs. This vignette will 
introduce the different options you can control and how to pass them into the **salesforcer** 
functions you're already familiar with.

### The new control argument

This new feature can be seen in the `sf_create` (and many other functions) as 
`control=list(...)`. The dots mean that you can pass any number of controls directly 
into the function. For example, the following code will create a record, but prevent 
its creation from showing up in the Chatter feeds by setting the `DisableFeedTrackingHeader`. 

```{r auth, include = FALSE}
library(salesforcer)
token_path <- here::here("tests", "testthat", "salesforcer_token.rds")
suppressMessages(sf_auth(token = token_path, verbose = FALSE))
```

```{r sample-create}
new_contact <- c(FirstName = "Test", LastName = "Contact-Create")
record <- sf_create(new_contact,
                    object_name = "Contact",
                    DisableFeedTrackingHeader = list(disableFeedTracking = TRUE), 
                    api_type = "SOAP")
```

```{r, include = FALSE}
deleted_records <- sf_delete(record$id)
```

You will notice that the argument `DisableFeedTrackingHeader` can be included right into 
the function without any documentation existing for it in the `sf_create` function. 
This is because the dots (`...`) allow you to pass over a dozen different control 
parameters and that documentation would be tedious to create and maintain over multiple 
functions in the package. However, you will notice in the documentation entry for 
the `control` argument there is a link to a function called `sf_control` which you 
can use to directly to pass into `control` or simply to review its documentation of all the 
possible control parameters and their defaults. This is where you can review the various 
control options in more detail before trying to set them.

You may have also noticed that the argument <a href="https://developer.salesforce.com/docs/atlas.en-us.api.meta/api/sforce_api_header_disablefeedtracking.htm" target="_blank">DisableFeedTrackingHeader</a> was formatted 
as a list with an element inside called `disableFeedTracking` set to `TRUE`. This may 
seem redundant but there are two reasons for this. First, this is exactly how the 
Salesforce APIs documents these options, which are typically referred to as "headers" 
because they are passed as a named header of the HTTP request and then the header fields 
and values are provided for that header. Second, some headers have multiple fields and values 
so a list is the only way to provide multiple named fields and values under a single header entity. 
For example, the <a href="https://developer.salesforce.com/docs/atlas.en-us.api.meta/api/sforce_api_header_duplicateruleheader.htm" target="_blank">DuplicateRuleHeader</a>, which controls whether 
the duplicate rules can be overridden when inserting records from the API, has three 
fields: `allowSave`, `includeRecordDetails`, and `runAsCurrentUser`. Supplying all 
three requires a list-like structure, which may seem redundant in other cases, but is 
necessary to follow.

```{r sample-create-w-duplicate}
new_contact <- c(FirstName = "Test", LastName = "Contact-Create")
record <- sf_create(new_contact,
                    object_name = "Contact",
                    DuplicateRuleHeader = list(allowSave = TRUE, 
                                               includeRecordDetails = FALSE, 
                                               runAsCurrentUser = TRUE))
```

```{r, include = FALSE}
deleted_records <- sf_delete(record$id)
```

Finally, you will notice in the example call that the `api_type` argument is set to "SOAP". 
This is because the `DisableFeedTrackingHeader` is a control that is only available when 
making calls via the SOAP API. You will receive a warning when trying to set control 
parameters for APIs or operations that do not recognize that control. For example, 
the following code tries to set the `BatchRetryHeader` for a call to the SOAP 
API which does not acknowledge that control. That control is only used with the Bulk 
1.0 API since its records as submitted in batches and automatic retry can be controlled.

```{r sample-create-w-warning}
new_contact <- c(FirstName = "Test", LastName = "Contact-Create")
record <- sf_create(new_contact,
                    object_name = "Contact",
                    BatchRetryHeader = list(`Sforce-Disable-Batch-Retry` = FALSE), 
                    api_type = "SOAP")
```

```{r, include = FALSE}
deleted_records <- sf_delete(record$id)
```

### Creating the control argument with sf_control

If this type of control structure is new to you, take a look at the documentation for 
the `glm` and `glm.control` functions. The way these two functions behave is exactly how 
functions like `sf_create` and `sf_control` work with each other. As demonstrated above 
you can pass any number of arbitrary controls into the function and they are all 
gathered up into the control by `control = list(...)`. However, you can specify the 
control directly like this: 

```{r sample-query}
new_contact <- c(FirstName = "Test", LastName = "Contact-Create")
records <- sf_query("SELECT Id, Name FROM Account LIMIT 1000",
                    object_name = "Account",
                    control = sf_control(QueryOptions = list(batchSize = 100)), 
                    api_type = "SOAP")
```

### Backwards compatibility for all_or_none and other named arguments

You may already be taking advantage of the `all_or_none` or `line_ending` arguments 
which are control arguments that were explicity included in functions. These argument 
essentially hard coded values to pass the `AllOrNoneHeader` and `LineEndingHeader` 
control parameters. Starting with the 0.1.3 release it is no longer necessary and 
preferable not to have an argument like `all_or_none` listed explicity as an argument 
since it can be provided in the `control` argument. Note: the `all_or_none` argument 
and other explicit control arguments will still be available in **salesforcer 0.1.3** 
but will provide a deprecated warning. They will be removed in the next CRAN release 
of the package so it will be important to update your code now if you are explicitly 
passing these arguments and see a deprecation warning.
  
### Reference Links

Below is a list of links that go directly to the control arguments (a.k.a headers) 
for the different APIs. I highly recommend reading this documentation before setting 
a control parameter in R so you know exactly what the behavior will be and how to 
specify it in R. You may notice that some controls are not included in the R package. 
Some may be added in the future if requested and some will not be added given the 
scope of the package. One final note is that some arguments in the REST API, like the 
"All or None" behavior is not a header, but a parameter in the API call. For this reason 
you will not see it listed in the REST API Headers section, but it is set in this R package 
using the `AllOrNoneHeader` argument in `sf_control` just to provide consistency between 
the SOAP and REST APIs. It would be confusing to have two arguments named differently, 
one for each API, but to do the exact same thing from R. For this reason, many of the 
control arguments match exactly as they are listed in the SOAP API, but can be used 
across other APIs even if not exactly written that way in the Salesforce documentation 
referenced below.

 * **SOAP API Headers**: <a href="https://developer.salesforce.com/docs/atlas.en-us.api.meta/api/soap_headers.htm" target="_blank">https://developer.salesforce.com/docs/atlas.en-us.api.meta/api/soap_headers.htm</a>
 * **REST API Headers**: <a href="https://developer.salesforce.com/docs/atlas.en-us.api_rest.meta/api_rest/headers.htm" target="_blank">https://developer.salesforce.com/docs/atlas.en-us.api_rest.meta/api_rest/headers.htm</a>
 * **Bulk 1.0 API Headers**: <a href="https://developer.salesforce.com/docs/atlas.en-us.api_asynch.meta/api_asynch/async_api_headers.htm" target="_blank">https://developer.salesforce.com/docs/atlas.en-us.api_asynch.meta/api_asynch/async_api_headers.htm</a>
 * **Bulk 2.0 API Headers**: None
 * **Metadata API Headers**: <a href="https://developer.salesforce.com/docs/atlas.en-us.api_meta.meta/api_meta/meta_headers.htm" target="_blank">https://developer.salesforce.com/docs/atlas.en-us.api_meta.meta/api_meta/meta_headers.htm</a>
