---
title: "Transitioning from RForcecom"
author: "Steven M. Mortimer"
date: "2021-07-03"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 4
    keep_md: true
vignette: >
  %\VignetteIndexEntry{Transitioning from RForcecom}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo = FALSE}
NOT_CRAN <- identical(tolower(Sys.getenv("NOT_CRAN")), "true")
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  purl = NOT_CRAN,
  eval = NOT_CRAN
)
options(tibble.print_min = 5L, tibble.print_max = 5L)
```

While writing the {salesforcer} package we were keenly aware that many folks 
were already using the {RForcecom} package to connect to Salesforce. In order 
to foster adoption and switching between the packages {salesforcer} replicates 
the functionality of many {RForcecom} functions so that you will only need to swap 
out `library(RForcecom)` for `library(salesforcer)` and still have 
production scripts perform as expected.

## RForcecom Removed from CRAN

As of June 9, 2021, the {RForcecom} package was removed from CRAN. You can still 
use it by installing from the archive, but we strongly recommend using 
{salesforcer} instead. The existing functionality in {RForcecom} has been further 
optimized within {salesforcer} and new functionality has been added too.

## Authentication

{salesforcer} supports OAuth 2.0 authentication which is preferred, but for 
backward compatibility provides the username-password authentication routine 
implemented by {RForcecom}. Here is an example running the function from 
each of the packages side-by-side and producing the same result.

```{r auth, eval=FALSE}
library(dplyr, warn.conflicts = FALSE)
library(salesforcer)
```

```{r auth-background, include = FALSE}
library(dplyr, warn.conflicts = FALSE)
library(salesforcer)

username <- Sys.getenv("SALESFORCER_USERNAME")
password <- Sys.getenv("SALESFORCER_PASSWORD")
security_token <- Sys.getenv("SALESFORCER_SECURITY_TOKEN")

sf_auth(username = username,
        password = password,
        security_token = security_token)
```

```{r, warning=FALSE}
# the RForcecom way
# RForcecom::rforcecom.login(username, paste0(password, security_token), 
#                            apiVersion=getOption("salesforcer.api_version"))
# replicated in salesforcer package
session <- salesforcer::rforcecom.login(username, 
                                         paste0(password, security_token), 
                                         apiVersion = getOption("salesforcer.api_version"))
session['sessionID'] <- "{MASKED}"
session
```

```{r, include=FALSE}
# keep using the session, just rename it to "session"
session <- salesforcer::rforcecom.login(username, 
                                        paste0(password, security_token), 
                                        apiVersion = getOption("salesforcer.api_version"))
```

Note that we must set the API version here because calls to session will not
create a new sessionId and then we are stuck with version 35.0 (the default from
`RForcecom::rforcecom.login()`). Some functions in {salesforcer} implement API
calls that are only available after version 35.0.

## CRUD Operations

"CRUD" operations (Create, Retrieve, Update, Delete) in the {RForcecom} package
only operate on one record at a time. One benefit to using the {salesforcer}
package is that these operations will accept a named vector (one record) or an
entire `data.frame` or `tbl_df` of records to churn through. However, rest
assured that the replicated functions behave exactly the same way if you are
hesitant to making the switch.

```{r, warning=FALSE}
object <- "Contact"
fields <- c(FirstName="Test", LastName="Contact-Create-Compatibility")

# the RForcecom way
# RForcecom::rforcecom.create(session, objectName=object, fields)

# replicated in salesforcer package
result <- salesforcer::rforcecom.create(session, objectName=object, fields)
result
```

Here is an example showing the reduction in code of using {salesforcer} if you 
would like to create multiple records.

```{r, warning=FALSE}
n <- 2
new_contacts <- tibble(FirstName = rep("Test", n),
                       LastName = paste0("Contact-Create-", 1:n))

# the RForcecom way (requires a loop)
# rforcecom_results <- NULL
# for(i in 1:nrow(new_contacts)){
#   temp <- RForcecom::rforcecom.create(session, 
#                                       objectName = "Contact", 
#                                       fields = unlist(slice(new_contacts,i)))
#   rforcecom_results <- bind_rows(rforcecom_results, temp)
# }

# the better way in salesforcer to do multiple records
salesforcer_results <- sf_create(new_contacts, object_name="Contact")
salesforcer_results
```

## Query

{salesforcer} also has better printing and type-casting when returning query
result thanks to features of the {readr} package.

```{r, warning=FALSE}
this_soql <- "SELECT Id, Email FROM Contact LIMIT 5"

# the RForcecom way
# RForcecom::rforcecom.query(session, soqlQuery = this_soql)

# replicated in salesforcer package
result <- salesforcer::rforcecom.query(session, soqlQuery = this_soql)
result

# the better way in salesforcer to query
salesforcer_results <- sf_query(this_soql)
salesforcer_results
```

## Describe

The {RForcecom} package has the function `rforcecom.getObjectDescription()`
which returns a `data.frame` with one row per field on an object. The same
function in {salesforcer} is named `sf_describe_object_fields()` and also has
better printing and datatype casting by using tibbles.

```{r, warning=FALSE}
# the RForcecom way
# RForcecom::rforcecom.getObjectDescription(session, objectName='Account')

# backwards compatible in the salesforcer package
result <- salesforcer::rforcecom.getObjectDescription(session, objectName='Account')

# the better way in salesforcer to get object fields
result2 <- salesforcer::sf_describe_object_fields('Account')
result2
```
