---
title: "Transitioning from RForcecom"
author: "Steven M. Mortimer"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 4
    keep_md: true
vignette: >
  %\VignetteIndexEntry{Getting Started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo = FALSE}
NOT_CRAN <- identical(tolower(Sys.getenv("NOT_CRAN")), "true")
NOT_CRAN <- FALSE
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  purl = NOT_CRAN,
  eval = NOT_CRAN
)
```

While writing the **salesforcer** package we were keenly aware that many folks 
are already using the **RForcecom** package to connect to Salesforce. In order 
to foster adoption and switching between the packages **salesforcer** replicates 
the functionality of many **RForcecom** functions so that you will only need to swap 
out `library(RForcecom)` for `library(salesforcer)` and still have your production 
tested scripts perform as usual.

### Authentication

**salesforcer** supports OAuth 2.0 authentication which is preferred, but for 
backward compatibility provides the username-password authentication routine 
implemented by **RForcecom**. Here is an example running the function from 
each of the packages side-by-side and producing the same result.

```{r auth, include = FALSE}
library(dplyr)
library(salesforcer)
## I grab the settings from the testing directory because that's where it is to be
## found on Travis
salesforcer_test_settings <- readRDS(file.path("..", 
                                               "tests", 
                                               "testthat",
                                               "salesforcer_test_settings.rds"))
username <- salesforcer_test_settings$username
password <- salesforcer_test_settings$password
security_token <- salesforcer_test_settings$security_token
```

```{r, warning=FALSE}
suppressWarnings(suppressMessages(library(dplyr)))
library(salesforcer)

# the RForcecom way
session1 <- RForcecom::rforcecom.login(username, paste0(password, security_token), 
                                       apiVersion=getOption("salesforcer.api_version"))
session1

# replicated in salesforcer package
session2 <- salesforcer::rforcecom.login(username, paste0(password, security_token), 
                                         apiVersion=getOption("salesforcer.api_version"))
session2
```

```{r, include=FALSE}
# keep using the session, just rename it to "session"
session <- session2
```

Note that we must set the API version here because  calls to session will not create 
a new sessionId and then we are stuck with version 35.0 (the default from 
RForcecom::rforcecom.login). Some functions in **salesforcer** implement API calls 
that are only available after version 35.0.

### CRUD Operations

"CRUD" operations (Create, Retrieve, Update, Delete) in the **RForcecom** package 
only operate on one record at a time. One benefit to using the **salesforcer** package 
is that these operations will accept a named vector (one record) or an entire `data.frame`
or `tbl_df` of records to churn through. However, rest assured that the replicated 
functions behave exactly the same way if you are hesitant to making the switch.

```{r, warning=FALSE}
object <- "Contact"
fields <- c(FirstName="Test", LastName="Contact-Create-Compatibility")

# the RForcecom way
result1 <- RForcecom::rforcecom.create(session, objectName=object, fields)
result1

# replicated in salesforcer package
result2 <- salesforcer::rforcecom.create(session, objectName=object, fields)
result2
```

Here is an example showing the reduction in code of using **salesforcer** if you 
would like to create multiple records.

```{r, warning=FALSE}
n <- 2
new_contacts <- tibble(FirstName = rep("Test", n),
                       LastName = paste0("Contact-Create-", 1:n))
rforcecom_results <- NULL
# the RForcecom way
for(i in 1:nrow(new_contacts)){
  temp <- RForcecom::rforcecom.create(session, 
                                      objectName = "Contact", 
                                      fields = unlist(slice(new_contacts,i)))
  rforcecom_results <- bind_rows(rforcecom_results, temp)
}
rforcecom_results

# the better way in salesforcer to do multiple records
salesforcer_results <- sf_create(new_contacts, "Contact")
salesforcer_results
```

### Query

**salesforcer** also has better printing and type-casting when returning query result
thanks to features of the readr package.

```{r, warning=FALSE}
this_soql <- "SELECT Id, Email FROM Contact"

# the RForcecom way
result1 <- RForcecom::rforcecom.query(session, soqlQuery = this_soql)
result1

# replicated in salesforcer package
result2 <- salesforcer::rforcecom.query(session, soqlQuery = this_soql)
result2

# the better way in salesforcer to do multiple records
salesforcer_results <- sf_query(this_soql)
salesforcer_results
```

In the future more features will be migrated from **RForcecom** to make the 
transition as seamless as possible.
