---
title: "Working with Attachments"
author: "Steven M. Mortimer"
date: "2020-06-07"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 4
    keep_md: true
vignette: >
  %\VignetteIndexEntry{Working with Attachments}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo = FALSE}
NOT_CRAN <- identical(tolower(Sys.getenv("NOT_CRAN")), "true")
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  purl = NOT_CRAN,
  eval = NOT_CRAN
)
```

### Attachments in Salesforce

Almost all records in Salesforce support attachments. Attachments are just blob data 
storage for an associated ParentId. A ParentId is the 18-character Salesforcer Id 
of the record that the attachment belongs to. To get started creating and updating 
attachments, first, load the **salesforcer** and **dplyr** packages and login, if needed.

```{r auth, include = FALSE}
suppressWarnings(suppressMessages(library(dplyr)))
library(here)
library(salesforcer)
token_path <- here::here("tests", "testthat", "salesforcer_token.rds")
suppressMessages(sf_auth(token = token_path, verbose = FALSE))
```

```{r load-package, eval=FALSE}
library(dplyr)
library(here)
library(salesforcer)
sf_auth()
```

The Attachment data, for example, the attachment's Id, ParentId, Name, Body, ModifiedDate, 
and other attributes are stored in the Attachment object, a Standard Object in Salesforce.

### Creating (Uploading) Attachments

Below we will cover 2 different methods of creating attachments: 

1. Uploading local files as Attachments one at a time using the SOAP and REST API (most common usage)
2. Uploading large batches of files using the Bulk API, which zips the files before upload

**Uploading local files as Attachments (SOAP and REST)**

When uploading an attachment stored locally (i.e. on your computer), you can provide 
an absolute or relative path to the current working directory. If the `Name` column 
is omitted, then the name of the attachment as it appears in Salesforce will be 
the same as the base file name and extension. For example, if the path provided is 
"/Documents/attachments/doc1.pdf", then the `Name` field will be set to "doc1.pdf".
In the sample code below we are leveraging the **here** package to consistently 
create relative paths. Although, this is not required, it is encouraged to increase 
the portability of your code. 

In the example below, we upload three attachments to two different parent records. 
Note: Make sure to replace the paths and ParentIds in the example below with file 
paths that exist on your local machine and Ids of records in your Salesforce org.

```{r}
# define the ParentIds where the attachments will be shown in Salesforce
parent_record_id1 <- "0016A0000035mJ4"
parent_record_id2 <- "0016A0000035mJ5"

# provide the file paths of where the attachments exist locally on your machine
# in this case we are referencing images included within the salesforcer package, 
# but any absolute or relative path locally will work
file_path1 <- system.file("extdata", "cloud.png", package="salesforcer")
file_path2 <- system.file("extdata", "logo.png", package="salesforcer")
file_path3 <- system.file("extdata", "old-logo.png", package="salesforcer")

# create a data.frame or tbl_df out of this information
attachment_details <- tibble(Body = rep(c(file_path1, 
                                          file_path2, 
                                          file_path3), 
                                        times=2),
                             ParentId = rep(c(parent_record_id1, 
                                              parent_record_id2), 
                                            each=3))

# create the attachments!
result <- sf_create_attachment(attachment_details)
result
```

### Downloading Attachments

After uploading attachments to Salesforce you can download them by first querying 
the metadata in the Attachment object. This metadata provides the Id for the blob 
data attachment for download. A convenience function, `sf_download_attachment()`, 
has been created to download attachments quickly. The example below shows how to 
query the metadata of attachments belonging to particular ParentId.

```{r}
# pull down all attachments associated with a particular record
queried_attachments <- sf_query(sprintf("SELECT Id, Body, Name, ParentId
                                         FROM Attachment
                                         WHERE ParentId IN ('%s', '%s')", 
                                         parent_record_id1, parent_record_id2))
queried_attachments
```

Before downloading the attachments using the Body it is important to consider 
whether the attachment names are repeated or duplicates. If so, then the attachments 
with the same exact name will be overwritten on the local filesystem as they are 
downloaded. To avoid this problem there are two common strategies:  

1. Create a new column (e.g. `unique_name`) that is the concatenation of the 
Attachment Id and the Attachment's name which is guaranteed to be unique.
2. Save the attachments in separate folders for each ParentId record.

As long as the same ParentId record doesn't name attachments with the same name, then 
Strategy #2 above will work. In addition, it may help better organize the documents 
if you are planning to download many and then upload again to Salesforce using the 
Bulk API as demonstrated later in this script.

```{r}
# create a new folder for each ParentId in the dataset
temp_dir <- tempdir()
for (pid in unique(queried_attachments$ParentId)){
  dir.create(file.path(temp_dir, pid), showWarnings = FALSE) # ignore if already exists
}

# create a new columns in the queried data so that we can pass this information 
# on to the function `sf_download_attachment()` that will actually perform the download
queried_attachments <- queried_attachments %>% 
  # Strategy 1: Unique file names (ununsed here, but shown as an example)
  mutate(unique_name = paste(Id, Name, sep='___')) %>% 
  # Strategy 2: Separate folders per parent
  mutate(Path = file.path(temp_dir, ParentId))

# download all of the attachments for a single ParentId record to their own folder
download_result <- mapply(sf_download_attachment, 
                          queried_attachments$Body, 
                          queried_attachments$Name, 
                          queried_attachments$Path)
download_result
```
```{r cleanup-1, include = FALSE}
sf_delete(queried_attachments$Id)
```

**Uploading large batches using the Bulk API**

The SOAP and REST APIs are good for working with a few attachements at a time. 
However, the Bulk API can be envoked using api_type="Bulk 1.0" to automatically take a 
`data.frame` or `tbl_df` of Attachment field data and create a ZIP file with CSV manifest 
that is required by that API to upload In the example above we downloaded the 3 attachments 
each belonging to 2 different parent records. Assuming that I have a directory on my computer

```{r}
# create the attachment metadata required (Name, Body, ParentId)
attachment_details <- queried_attachments %>% 
  mutate(Body = file.path(Path, Name)) %>% 
  select(Name, Body, ParentId) 
```
```{r}
result <- sf_create_attachment(attachment_details, api_type="Bulk 1.0")
result
```
```{r cleanup-2, include = FALSE}
for (pid in unique(queried_attachments$ParentId)){
  unlink(file.path(temp_dir, pid), recursive=TRUE) # remove directories...
}
sf_delete(result$Id) #... and records in Salesforce
```

### Extending to Documents and Other Blob Data

The commands for working with Attachments also work for uploading documents and 
other blob data as well. Documents are just like Attachments in Salesforce except 
instead of having an associated ParentId they have an associated FolderId where 
the blob will be associated with upon creation. Here is a brief example of uploading 
a PDF ("Document") to a Folder 

```{r}
# the function supports inserting all types of blob content, just update the 
# object_name argument to add the PDF as a Document instead of an Attachment
document_details <- tibble(Name = "Data Wrangling Cheatsheet - Test 1",
                           Description = "RStudio cheatsheet covering dplyr and tidyr.",
                           Body = system.file("extdata", 
                                              "data-wrangling-cheatsheet.pdf",
                                              package="salesforcer"),
                           FolderId = "00l6A000001EgIwQAK",
                           Keywords = "test,cheatsheet,document")
result <- sf_create_attachment(document_details, object_name = "Document")
result
```
```{r cleanup-3, include = FALSE}
sf_delete(result$id)
```

With Documents, users are also able to save storage by specifying a `Url` instead 
of a a file path where the `Body` content is stored locally. Specifying the `Url` 
field will reference the URL instead of uploading into the Salesforce org, thereby 
saving space if limited in your organization.

```{r}
cheatsheet_url <- "https://rstudio.com/wp-content/uploads/2015/02/data-wrangling-cheatsheet.pdf"
document_details <- tibble(Name = "Data Wrangling Cheatsheet - Test 2",
                           Description = "RStudio cheatsheet covering dplyr and tidyr.",
                           Url = cheatsheet_url,  
                           FolderId = "00l6A000001EgIwQAK",
                           Keywords = "test,cheatsheet,document")
result <- sf_create_attachment(document_details, object_name = "Document")
result
```
```{r cleanup-4, include = FALSE}
sf_delete(result$id)
```

### Reference Links

Below is a list of links to existing Salesforce documentation that provide more 
detail into how Attachments, Documents, and other blob data are handled via their 
APIs. As with many functions in **salesforcer**, we have tried to tranlate these 
functions exactly as they are described in the Salesforce documentation so that 
they are flexible enough to handle most all cases that the APIs were intended to 
support.

 * **Attachment Object**: <a href="https://developer.salesforce.com/docs/atlas.en-us.api.meta/api/sforce_api_objects_attachment.htm" target="_blank">https://developer.salesforce.com/docs/atlas.en-us.api.meta/api/sforce_api_objects_attachment.htm</a>
 
 * **Document Object**: <a href="https://developer.salesforce.com/docs/atlas.en-us.api.meta/api/sforce_api_objects_document.htm" target="_blank">https://developer.salesforce.com/docs/atlas.en-us.api.meta/api/sforce_api_objects_document.htm</a>
 
 * **REST API Upload Attachment**: <a href="https://developer.salesforce.com/docs/atlas.en-us.api_rest.meta/api_rest/dome_sobject_insert_update_blob.htm" target="_blank">https://developer.salesforce.com/docs/atlas.en-us.api_rest.meta/api_rest/dome_sobject_insert_update_blob.htm</a> 
 
 * **Bulk API Upload Attachments**: <a href="https://developer.salesforce.com/docs/atlas.en-us.api_asynch.meta/api_asynch/binary_intro.htm" target="_blank">https://developer.salesforce.com/docs/atlas.en-us.api_asynch.meta/api_asynch/binary_intro.htm</a>
