---
title: "Working with Attachments"
author: "Steven M. Mortimer"
date: "2019-07-26"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 4
    keep_md: true
vignette: >
  %\VignetteIndexEntry{Working with Attachments}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo = FALSE}
NOT_CRAN <- identical(tolower(Sys.getenv("NOT_CRAN")), "true")
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  purl = NOT_CRAN,
  eval = NOT_CRAN
)
```

### Creating an Attachment (Uploading it)

Almost all records in Salesforce support attachments. Attachments are just blob data 
storage for an associated ParentId. A ParentId is the 18-character Salesforcer Id 
of the record that the attachment belongs to. To get started creating and updating 
attachments, first, load the **salesforcer**, **dplyr**, and **purrr** packages 
and login, if needed.

```{r auth, include = FALSE}
suppressWarnings(suppressMessages(library(dplyr)))
library(salesforcer)
token_path <- here::here("tests", "testthat", "salesforcer_token.rds")
suppressMessages(sf_auth(token = token_path, verbose = FALSE))
```

```{r load-package, eval=FALSE}
suppressWarnings(suppressMessages(library(dplyr)))
suppressWarnings(suppressMessages(library(purrr)))
library(salesforcer)
sf_auth()
```

The attachment metadata like the attachment's Id, ParentId, Name, modified timestamp 
and other attributes are kept in the Attachment object, a Standard Object in Salesforce.

### Creating Attachments

 - Simple example (SOAP and REST)
 - Url based attachment bodies
 - Bulk API example

### Extending to Documents and Other Blob Data

The commands we have shown above work for uploading documents and other blob data as well. 
Documents are just like Attachments in Salesforce except instead of having an 
associated ParentId they have an associated FolderId where the blob will be associated 
with upon creation.
