---
title: "Working with the Bulk APIs"
author: "Steven M. Mortimer"
date: "2020-06-07"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 4
    keep_md: true
vignette: >
  %\VignetteIndexEntry{Working with the Bulk APIs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo = FALSE}
NOT_CRAN <- identical(tolower(Sys.getenv("NOT_CRAN")), "true")
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  purl = NOT_CRAN,
  eval = NOT_CRAN
)
```

### Using the Bulk API

First, load the **salesforcer** and **dplyr** packages and login, if needed. 

```{r auth, include = FALSE}
suppressWarnings(suppressMessages(library(dplyr)))
library(salesforcer)
token_path <- here::here("tests", "testthat", "salesforcer_token.rds")
suppressMessages(sf_auth(token = token_path, verbose = FALSE))
```

```{r load-package, eval=FALSE}
suppressWarnings(suppressMessages(library(dplyr)))
library(salesforcer)
sf_auth()
```

For really large inserts, updates, deletes, upserts, and queries you can just add 
`api_type = "Bulk 1.0"` to most functions to get the benefits of using the Bulk API 
instead of the SOAP or REST APIs. The change you have to make from using the REST 
API to the Bulk 1.0 API as simple as adding `api_type="Bulk 1.0"` to your function 
arguments. First, let's build a `tbl_df` with two new records to be created.

```{r}
n <- 2
prefix <- paste0("Bulk-", as.integer(runif(1,1,100000)), "-")
new_contacts1 <- tibble(FirstName = rep("Test", n),
                        LastName = paste0("Contact-Create-", 1:n),
                        My_External_Id__c=paste0(prefix, letters[1:n]))
new_contacts2 <- tibble(FirstName = rep("Test", n),
                        LastName = paste0("Contact-Create-", 1:n),
                        My_External_Id__c=paste0(prefix, letters[1:n]))
```

There are some differences in the way the REST API returns response information vs. 
the Bulk 1.0 API. However, the changes in Salesforce are exactly the same for these 
two calls.

```{r}
# REST
rest_created_records <- sf_create(new_contacts1, object_name="Contact", api_type="REST")
rest_created_records

# Bulk
bulk_created_records <- sf_create(new_contacts2, object_name="Contact", api_type="Bulk 1.0")
bulk_created_records
```

### A Complete Bulk API Workflow

To show a more lengthy example of using the Bulk 1.0 API, below is a workflow of 
that creates 2 records, queries them, and deletes them. This is just an example. 
Typically, you'd want to use the Bulk APIs over the REST or SOAP APIs when dealing 
with over 10,000 records.

```{r, include=FALSE}
n <- 2
prefix <- paste0("Bulk-", as.integer(runif(1,1,100000)), "-")
new_contacts <- tibble(FirstName = rep("Test", n),
                       LastName = paste0("Contact-Create-", 1:n),
                       My_External_Id__c=paste0(prefix, letters[1:n]))
```
```{r}
object <- "Contact"
created_records <- sf_create(new_contacts, object_name=object, api_type="Bulk 1.0")
created_records

# query bulk
my_soql <- sprintf("SELECT Id,
                           FirstName, 
                           LastName
                    FROM Contact 
                    WHERE Id in ('%s')", 
                   paste0(created_records$Id , collapse="','"))

queried_records <- sf_query(my_soql, object_name=object, api_type="Bulk 1.0")
queried_records

# delete bulk
deleted_records <- sf_delete(queried_records$Id, object_name=object, api_type="Bulk 1.0")
deleted_records
```

### Query Limitations

There is one limitation to Bulk queries is that it does not support the following 
operations or structures of SOQL: 

 - COUNT
 - ROLLUP
 - SUM
 - GROUP BY CUBE
 - OFFSET
 - Nested SOQL queries
 - Relationship fields
 - Compound address fields
 - Compound geolocation fields
 
### Using the Bulk 2.0 API
 
The **salesforcer** package also implements the Bulk 2.0 API which has better 
speed than the Bulk 1.0 API but sacrifices consistency in the ordering of the 
result records since they are batched and processed asynchronously. In the example 
below we first create 10 records in five batches of two records using both the 
Bulk 1.0 API and then the Bulk 2.0 API. After looking at the results you can see 
how the returned results are ordered when we process them using the Bulk 1.0 API 
because after the data is split into batches each batch is processed synchronously 
which preserves the order of the rows in the returned output. This is not necessarily 
true of the Bulk 2.0 API which is the point (it's asynchronicity).

**Bulk 1.0**

```{r, include=FALSE}
n <- 10
prefix <- paste0("Bulk-", as.integer(runif(1,1,100000)), "-")
new_contacts <- tibble(FirstName = rep("Test", n),
                       LastName = paste0("Contact-Create-", 1:n),
                       test_number__c = 1:10,
                       My_External_Id__c=paste0(prefix, letters[1:n]))
```
```{r, eval=FALSE}
n <- 10
new_contacts <- tibble(FirstName = rep("Test", n),
                       LastName = paste0("Contact-Create-", 1:n), 
                       test_number__c = 1:10)
```

```{r}
created_records_v1 <- sf_create(new_contacts, "Contact", api_type="Bulk 1.0", batch_size=2)
created_records_v1

# query the records so we can compare the ordering of the Id field to the 
# original dataset
my_soql <- sprintf("SELECT Id,
                           test_number__c
                    FROM Contact 
                    WHERE Id in ('%s')", 
                   paste0(created_records_v1$Id , collapse="','"))
queried_records <- sf_query(my_soql)
queried_records <- queried_records %>% 
  arrange(test_number__c)

# same ordering of rows!
cbind(created_records_v1 %>% select(Id), queried_records)
```

The Bulk 2.0 API returns every single field that was included in the call so if you 
have an identifying key your dataset, then it should not be a problem to join on that 
key with your original data to bring in the newly assigned Salesforce Id that is 
generated when the record is created in Salesforce. However, I have find it wasteful 
to transfer all of the field information back after the query and have not found 
a significant performance improvement between the Bulk 1.0 and Bulk 2.0. 

**Bulk 2.0**

```{r, include=FALSE}
n <- 10
prefix <- paste0("Bulk-", as.integer(runif(1,1,100000)), "-")
new_contacts <- tibble(FirstName = rep("Test", n),
                       LastName = paste0("Contact-Create-", 1:n),
                       test_number__c = 1:10,
                       My_External_Id__c=paste0(prefix, letters[1:n]))
created_records_v2 <- sf_create(new_contacts, "Contact", api_type="Bulk 2.0")
created_records_v2 <- created_records_v2 %>% 
  select(-My_External_Id__c)
```
```{r, eval=FALSE}
created_records_v2 <- sf_create(new_contacts, "Contact", api_type="Bulk 2.0", batch_size=2)
```

The result in Bulk 2.0 API returns all fields so that you can still associate the 
records if the ordering was changed during processing. However, in this simple 
case the ordering is still preserved because the number of records processed in 
this example was not enough to create separate batches to process asynchronously. 
Finally, note that the field names are different from the Bulk 1.0 API, this is 
the Salesforcer convention for the Bulk 2.0 API.

```{r}
created_records_v2
```

For these above I typically prefer using the Bulk 1.0 API when creating or updating 
records. That way I can be confident in that the order of the records returned 
from the process matches the order of the original dataset I put into the process. 
All of this does not matter when deleting records, so if you want minor performance 
improvements then switch to the Bulk 2.0 when deleting records.

```{r}
sf_delete(c(created_records_v1$Id, created_records_v2$sf__Id), 
          object_name = "Contact", api_type="Bulk 2.0")
```
